---
title: "od vs cc data"
author: "ciara"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document: 
    df_print: paged
    css: stylesheet.css
    number_sections: yes
    toc: yes
    toc_float: true
    toc_depth: 3
    code_folding: show
editor_options:
  chunk_output_type: console
---

```{r setup, echo = TRUE, message=FALSE, warning=FALSE}
# load libraries
library(tidyverse) 
library(latex2exp)
library(readxl)
library(growthcurver)
library(cowplot)

# source all relevant scripting files
source(file.path("scripts", "plotting_functions.R"))

# global knitting options for automatic saving of all plots as .png and .pdf
knitr::opts_chunk$set(
  dev = c("png", "pdf"), fig.keep = "all",
  dev.args = list(pdf = list(encoding = "WinAnsi", useDingbats = FALSE)),
  fig.path = file.path("fig_output", paste0(gsub("\\.[Rr]md", "", knitr::current_input()), "_"))
)
```

Data looking at paracoccus denitrificans in both chemostat (exps 1,2) and batch (exps 3,4) culture. Was looking to see if we can calibrate between coulter counter cells/mL and how the coulter counter does with fixed cells. Cells were fixed with 1% formaldehyde and stored for six months.

# load data
```{r}
df <- read_excel("data/chemostat_ccvsod.xlsx")

```

# convert to hours
```{r}

df_time <- df %>% 
  # introduce time in hours, sorted by experiment. 
  group_by(exp) %>% 
  mutate(time.hours = (time - min(time)) %>% as.numeric("hours") ) %>% 
  ungroup()

```

# OD vs CC
## do some calcs
```{r}
# od:cc, cells/mL * cell volume, percent of fixed to fresh counts

df_time <- df_time %>% 
  mutate(
    ratio = (cc/1e9)/od,
    `1e9 cells/od` = ratio,
    `1e9 cells * size [µm] ^3` = cc/1e9 * size^3,
    percent_og_c = (fix_cc/cc) * 100,
    percent_og_s = (fix_size/ size) *100
  )

```

## OD, cc, ratio vs time

```{r}
  df_time %>% 
  filter(exp > 2) %>%
  #filter(exp <=2) %>% 
  gather("var", "val", od, size, cc,`1e9 cells * size [µm] ^3`, `1e9 cells/od`) %>% 
  ggplot() +
  aes(time.hours, val, color = as.factor(exp)) +
  facet_grid(var ~., scales = "free_y")+
  geom_point(size = 2)


ggsave("fig_output/batch_summary.pdf", width = 6, height = 9, units = c("in"))
```


## plot OD vs cc
```{r}
#chemostats = exps 1 and 2

df_time %>% 
  #filter(exp == 4) %>% 
  #filter(exp <= 2) %>% 
  ggplot() +
  aes(od, cc, color = as.factor(exp)) +
  geom_point() 

```

## cc vs time for batch 4
```{r}
df_time %>% 
  filter(exp == 4) %>% 
  ggplot() +
  aes(time.hours, cc, color = as.factor(rep)) +
  geom_point()

# yikes 
```

## plot ratio (od/cc) vs time
```{r}
df_time %>% 
  #filter(temp == 30) %>% 
  ggplot() +
  aes(time.hours, ratio, color = as.factor(exp)) +
  geom_point()
```

## plot vs size
```{r}
df_time %>% 
  #filter(exp > 2) %>% 
  ggplot() +
  aes(time.hours, size, color = as.factor(exp)) +
  geom_point()
```

# orginal vs fixed
## original counts vs fixed
```{r}
fixcount_p <- df_time %>% 
  filter(exp < 3) %>% 
  ggplot() +
  aes(cc, fix_cc, color = as.factor(exp)) +
  geom_abline() +
  geom_point()

fixcount_p
```

## size vs fixed cells
```{r}
fixsize_p <- df_time %>% 
  ggplot() +
  aes(size, fix_size, color = as.factor(exp)) +
  geom_abline() +
  geom_point()

fixsize_p
# some of those fixes sizes are skewed by weird tails that appeared in the fixed cell counts, excluding them gets them much closer to the orginal size
```

# fixed cells number vs. size ratio

```{r}
df_time %>% 
  mutate(
    `fix cc/cc` = fix_cc/cc,
    `fix size/size` = fix_size/size
  ) %>% 
  ggplot() +
  aes(`fix size/size`, `fix cc/cc`, color = factor(exp)) +
  geom_point(data = tibble(x = 1, y = 1), map = aes(x, y, color = NULL), shape = 17, size = 8) +
  geom_point(size = 4)

ggsave("fig_output/fixedvsfresh_summary.pdf")

```


## percent change
```{r}
df_time %>% 
  ggplot() +
  aes(exp, percent_og_c) +
  geom_point()
```

